LAGOS lake depth

### 2021-08-04 figures/01_contrasts.Rmd
linting

```{r define-plots, echo=FALSE}
contrast_plot <- function(..., plot_type){
  theme_opts <- cowplot::theme_cowplot()

  if(plot_type == "violin"){
    violin_plot(...) + theme_opts
  }else{
    pointrange_plot(...) + theme_opts
  }
}

violin_plot <- function(x, grp, varb, label){
  ggplot(data = x) +
  geom_violin(aes(x = UQ(rlang::sym(grp)), y = UQ(rlang::sym(varb))), 
              adjust = 0.8, scale = "count", 
              draw_quantiles = c(0.25, 0.5, 0.75)) +
    ylab(label) + xlab("")
}

quantile_collapse <- function(x){
  list(tibble::enframe(
    quantile(x, probs = c(0.05,0.5,0.95), na.rm = TRUE)))
}

pointrange_plot <- function(x, grp, varb, label){
  x %>%
  group_by(UQ(rlang::sym(grp))) %>%
    dplyr::select(UQ(rlang::sym(grp)), UQ(rlang::sym(varb))) %>%
    summarize_all(quantile_collapse) %>%
    unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = UQ(rlang::sym(grp)), y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab(label) + xlab("")
}
```

@@

```{r gg_depth_connectivity, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_depth_conny_class <- dt %>%
  dplyr::filter(!is.na(lake_connectivity_class)) %>%
  group_by(lake_connectivity_class) %>%
  summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = lake_connectivity_class, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab("")
```

@@

gg_depth_glaciated_class <- dt %>%
  dplyr::filter(!is.na(lake_glaciatedlatewisc)) %>%
  group_by(lake_glaciatedlatewisc) %>%
  summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = lake_glaciatedlatewisc, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab("")
```

@@

(gg_depth_xtabs <- group_by(dt_sub, 
                           lake_glaciatedlatewisc, reservoir_class) %>%
    summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  tidyr::unite(cat, lake_glaciatedlatewisc:reservoir_class) %>%
  ggplot() +
geom_pointrange(aes(x = cat, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab(""))

```

@@

```{r gg_tally_waterareaclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# panel histograms that tally number of lakes with max depth by waterarea_class
# include sum across all types
gg_tally_waterareaclass <- read.csv("../data/lagosus_depth_predictors.csv", 
               stringsAsFactors = FALSE) %>%
  mutate(area_class =
           smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 300, 400,
                                               1000, 20000, Inf),
             labels = ~paste(sep="-", thousand_k(.y[1]), thousand_k(.y[2])))
         ) %>%
  group_by(area_class) %>%
  tally() %>%
  ggplot(aes(x = area_class, y = n)) +
  geom_col() + 
  geom_text(aes(label = n), vjust = -0.5) +
  ylim(0, 11000) + 
  xlab("Lake area class (ha)")
```

>>>>

```{r define-plots, echo=FALSE}
contrast_plot <- function(..., plot_type) {
  theme_opts <- cowplot::theme_cowplot()

  if (plot_type == "violin") {
    violin_plot(...) + theme_opts
  } else {
    pointrange_plot(...) + theme_opts
  }
}

violin_plot <- function(x, grp, varb, label) {
  ggplot(data = x) +
    geom_violin(aes(x = UQ(rlang::sym(grp)), y = UQ(rlang::sym(varb))),
      adjust = 0.8, scale = "count",
      draw_quantiles = c(0.25, 0.5, 0.75)) +
    ylab(label) + xlab("")
}

quantile_collapse <- function(x) {
  list(tibble::enframe(
    quantile(x, probs = c(0.05, 0.5, 0.95), na.rm = TRUE)))
}

pointrange_plot <- function(x, grp, varb, label) {
  x %>%
    group_by(UQ(rlang::sym(grp))) %>%
    dplyr::select(UQ(rlang::sym(grp)), UQ(rlang::sym(varb))) %>%
    summarize_all(quantile_collapse) %>%
    unnest() %>%
    tidyr::spread(name, value) %>%
    ggplot() +
    geom_pointrange(aes(x = UQ(rlang::sym(grp)), y = `50%`,
      ymin = `5%`, ymax = `95%`)) +
    ylab(label) + xlab("")
}
```

@@

```{r gg_depth_connectivity, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
gg_depth_conny_class <- dt %>%
  dplyr::filter(!is.na(lake_connectivity_class)) %>%
  group_by(lake_connectivity_class) %>%
  summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05, 0.5, 0.95), na.rm = TRUE)))) %>%
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
  geom_pointrange(aes(x = lake_connectivity_class, y = `50%`,
    ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab("")
```

@@

gg_depth_glaciated_class <- dt %>%
  dplyr::filter(!is.na(lake_glaciatedlatewisc)) %>%
  group_by(lake_glaciatedlatewisc) %>%
  summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05, 0.5, 0.95), na.rm = TRUE)))) %>%
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
  geom_pointrange(aes(x = lake_glaciatedlatewisc, y = `50%`,
    ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab("")
```

@@

(gg_depth_xtabs <- group_by(dt_sub,
  lake_glaciatedlatewisc, reservoir_class) %>%
  summarize(lake_maxdepth_m = list(tibble::enframe(
    quantile(lake_maxdepth_m, probs = c(0.05, 0.5, 0.95), na.rm = TRUE)))) %>%
  unnest() %>%
  tidyr::spread(name, value) %>%
  tidyr::unite(cat, lake_glaciatedlatewisc:reservoir_class) %>%
  ggplot() +
  geom_pointrange(aes(x = cat, y = `50%`,
    ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab(""))
```

@@

```{r gg_tally_waterareaclass, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# panel histograms that tally number of lakes with max depth by waterarea_class
# include sum across all types
gg_tally_waterareaclass <- read.csv("../data/lagosus_depth_predictors.csv",
  stringsAsFactors = FALSE) %>%
  mutate(area_class =
    smart_cut(lake_waterarea_ha, c(1, 4, 40, 80, 300, 400,
      1000, 20000, Inf),
    labels = ~ paste(sep = "-", thousand_k(.y[1]), thousand_k(.y[2])))
  ) %>%
  group_by(area_class) %>%
  tally() %>%
  ggplot(aes(x = area_class, y = n)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.5) +
  ylim(0, 11000) +
  xlab("Lake area class (ha)")
```
